A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-09-18 doc/large_strain.tex
more citations

The performance of finite element (FEM) solvers on modern computer architectures is typically memory bound.
The main cause for this is that loading matrix elements is significantly slower than performing the arithmetic operations when solving the problem.
In order to improve the performance of the iterative solver, so-called matrix-free methods are
widely adopted in fluid mechanics community, where the matrix vector products are formed on-the-fly \cite{kronbichler12,May2015, Krank2017}. Such methods can also be efficiently implemented on GPUs \cite{Abdelfattah2016}.

To the best of our knowledge, the matrix-free methods are not widely adopted within the solid mechanics community.
A possible reason is that tangent operators (coming from linearization of the nonlinear balance equations) are
more elaborate as compared to the fluid mechanics and it is not obvious whether the matrix-free methods can be advantageous in this case. Another reason could be the smoothness of the underlying solution. FEM with linear or quadratic elements (to avoid locking) is often adopted in solid mechanics community. However we note that large scale computations with $10^{12}$ unknowns even with linear FEM inevitably require matrix-free approach as there is simply not enough memory to store the sparse tangent matrix \cite{Gmeiner2016}. Therefore we believe that for large scale computations in solid mechanics it is crucial to adopt matrix-free approaches and develop efficient multigrid solvers.

>>>>

The performance of finite element (FEM) solvers on modern computer architectures is typically memory bound.
The main cause for this is that loading matrix elements is significantly slower than performing the arithmetic operations when solving the problem.
In order to improve the performance of the iterative solver, so-called matrix-free methods are
widely adopted in fluid mechanics community, where the matrix vector products are formed on-the-fly \cite{kronbichler12,May2015, Krank2017, Brown2010}. Such methods can also be efficiently implemented on GPUs \cite{Abdelfattah2016}.

To the best of our knowledge, the matrix-free methods are not widely adopted within the solid mechanics community.
We are only aware of a single paper in preparation that deals with small strain linear elasticity \cite{Clevenger2018}.
A possible reason is that tangent operators (coming from linearization of the nonlinear balance equations) are
more elaborate as compared to the fluid mechanics and it is not obvious whether the matrix-free methods can be advantageous in this case. Another reason could be the smoothness of the underlying solution. FEM with linear or quadratic elements (to avoid locking) is often adopted in solid mechanics community. However we note that large scale computations with $10^{12}$ unknowns even with linear FEM inevitably require matrix-free approach as there is simply not enough memory to store the sparse tangent matrix \cite{Gmeiner2016}. Therefore we believe that for large scale computations in solid mechanics it is crucial to adopt matrix-free approaches and develop efficient multigrid solvers.
